---
title: "Homework Tidyverse"
subtitle: "Hamid Semiyari"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.height = 3,
  fig.width = 6,
  fig.align = "center"
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
#| warning: false
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, gapminder, broom)
```

# Problem 1:

Copy paste and run the tribble given below.

```{r}
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
) -> dt
```

## 1_a:

- Use and show a `map` function to find the "mean" of each column of the `dt` data table
```{r}
column_means <- map_dbl(dt, mean)
print(column_means)
```


## 1_b:

- Use and show a `map` function to find the "standard deviation" of each column of the `dt` data table.
```{r}
column_std <- map_dbl(dt, sd)
column_std
```

## 1_c:

- Use and show a map function that will calculate the "square root" of each value of each column of the data table `dt`.
```{r}
column_square_root <- dt |>
  map_df(sqrt)
print(column_square_root)
```


## 1_d:

- Use R code to find the "mean", "max", "1st Quartile", "3rd Quartile", "Median", and "Min" for each column of the `dt` data table. (Hint: You do not have to use a map function)

```{r}
summary(dt)
```


# Problem 2:

Write a function that uses a for loop that, for each iteration, randomly draws 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp()`) and calculates its "mean". It should do this 10,000 times. Choose an appropriate plot to plot the distribution of "means".

- Solution: We want to fill a vector with values. We should create this vector beforehand using the `vector()` function.

- We want to create a vector of size 10 and we want its values to be numeric.

- Then we want to randomly draw 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp(5, rate = 1)`) and calculates its "mean"

- Now we want to create a vector that has length of 10 and its values are mean of this randomly selected exponential functions

- Now we want to draw the histogram. But we need to have a data frame. Let us define a data frame.

- Now we can draw a histogram

  - **Now, let's solve the problem all in one step**
  
```{r}
sim_means <- vector(mode = 'numeric', length = 10000)
for (i in 1:10000) {
  #Randomly draws 5 observations with a rate of 1
  random_draws <- rexp(5, rate = 1)
  
  #Calculating the mean of those 5 observations
  sim_means[i] <- mean(random_draws)
}

#converting the results into a data frame
df_means <- data.frame(mean_values = sim_means)
  
  #creating a histogram from the data frame
ggplot(df_means, aes(x = mean_values)) +
    geom_histogram(fill = 'lightgreen', colour = 'pink', bins = 30) + 
    labs(
    title = "Distribution of 10,000 Sample Means",
    x = "Sample Mean",
    y = "Count"
  ) +
    theme_minimal()
```
  

## 2_a:

- Repeat part 1 by using a `map_*()` function.

```{r}
# 1. Run the simulation 10,000 times using map_dbl
sim_means_map <- map_dbl(1:10000, ~ mean(rexp(5, rate = 1)))

# 2. Convert the vector to a data frame for ggplot2
df_means_map <- data.frame(mean_values = sim_means_map)

# 3. Generate the histogram plot
ggplot(df_means_map, aes(x = mean_values)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(
    title = "Distribution of 10,000 Sample Means (Using map_dbl)",
    x = "Sample Mean",
    y = "Count"
  ) +
  theme_minimal()
```

## 2_b:

- Repeat part 1 by using the `replicate()` function.

```{r}
# 1. Run the simulation 10,000 times using replicate
sim_means_rep <- replicate(10000, mean(rexp(5, rate = 1)))

# 2. Convert the vector to a data frame for ggplot2
df_means_rep <- data.frame(mean_values = sim_means_rep)

# 3. Generate the histogram plot
ggplot(df_means_rep, aes(x = mean_values)) +
  geom_histogram(fill = "lightpink", color = "black", bins = 30) +
  labs(
    title = "Distribution of 10,000 Sample Means (Using replicate)",
    x = "Sample Mean",
    y = "Count"
  ) +
  theme_minimal()
```


## 2_c:

- Use a another for loop that will print out plots for sample sizes of 5, 10, and 20 observations (instead of just 5).

```{r}
sample_sizes <- c(5, 10, 20)
# Run a loop where 'n' takes the value of 5, then 10, then 20
for (n in sample_sizes) {
  
  # 1. Run the simulation 10,000 times using the current sample size 'n'
  sim_means <- replicate(10000, mean(rexp(n, rate = 1)))
  
  # 2. Put the vector into a temporary data frame for plotting
  df_temp <- data.frame(mean_values = sim_means)
  
  # 3. Create the histogram plot
  p <- ggplot(df_temp, aes(x = mean_values)) +
    geom_histogram(fill = "plum", color = "black", bins = 30) +
    labs(
      title = paste("Distribution of 10,000 Sample Means (n =", n, ")"),
      x = "Sample Mean",
      y = "Count"
    ) +
    theme_minimal()
  
  # 4. CRITICAL: Inside a for loop, you must explicitly print() the plot
  print(p)
}
```


# Problem 3:

- Use and show R coding to calculate the "standard deviation" for each variable of the data table `mtcars` using the “Special For Loop Method”.

**Note:**

- There are two methods. Please see a quick explanation for both methods in below:

  - Method 1:

```{r}
  multi_return <- function() {
    my_list <- list("color" = "red", "size" = 20, "shape" = "round")
    return(my_list) 
  }
```

```{r}
a <- multi_return()
a$color

a<- multi_return()
a$shape

a<- multi_return()
a$size
```

- Method 2. The "Special For Loop Method"

```{r}
#| eval: false
output <- vector("double", ncol(df))  # 1. output (use appropriate data frame instead of df)
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output
```

- **Solution:** coding to calculate the "standard deviation" for each variable

```{r}
sd_results <- vector(mode = "numeric", length = ncol(mtcars))

# 2. Assign names to the vector positions so you know which column is which
names(sd_results) <- colnames(mtcars)

# 3. Run the for loop over every column index
for (i in 1:ncol(mtcars)) {
  # Calculate the standard deviation for column 'i' and save it in position 'i'
  sd_results[i] <- sd(mtcars[[i]])
}

# 4. Print the final calculated values
print(sd_results)
```

# Problem 4:

Generate a sequence of dates from **January 1, 2015** to **December 31, 2025**, spaced by every **two months**. Extract the **year**, **quarter**, and **ISO week number** for each date.

```{r}
library(lubridate)

date_seq <- seq(from = as.Date("2015-01-01"),
                to = as.Date("2025-12-31"),
                by = "2 months")

date_breakdown <- data.frame(
  Original_Date = date_seq,
  Year = year(date_seq),
  Quarter = quarter(date_seq),
  ISO_Week = isoweek(date_seq)
)

head(date_breakdown, 10)
```


# Problem 5

Given the following dates, compute the difference in **months** and **weeks** between each consecutive pair.

``` r
sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")
```

# Problem 6

Given a list of mixed date formats, use `map()` and `possibly()` from `purrr` to safely convert them to **Date** format and extract the **month name**.

``` r
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")
```
